---
title: "POLS 299-F06: Religion in American Politics"
format:
    html: default
    pdf: 
      documentclass: article
      fontsize: 12pt
      colorlinks: true
      citecolor: DarkViolet
      linkcolor: DarkViolet
      urlcolor: DarkViolet
      geometry: margin=1in
      output-file: "POLS299-F06_syllabus"
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tibble)
library(pander)
```

::: {.content-visible when-format="pdf"}

\begin{center}

\vspace{-5em}

\url{ {{< var course.url >}} } \vspace{-1ex}

{{< var course.time1 >}} {{< var course.days >}} \vspace{-1ex}

{{< var course.time2 >}} {{< var course.days >}} \vspace{-1ex}

{{< var course.location >}} \vspace{-1ex}

{{< var course.semester >}}

\end{center}


## Instructor

{{< var instructor.name >}} \vspace{-1ex}

Email: [{{< var instructor.email >}}](mailto:{{< var instructor.email >}}) \vspace{-1ex}

Office: {{< var instructor.office >}} \vspace{-1ex}

Office hours: [M 9-11 am and Th 3:30-5:30 pm]({{< var instructor.appointment_url >}}) and by appointment

:::


::: {.content-visible when-format="html"}

::: {.grid .course-details .course-details-smaller}

::: {.g-col-6}
### Course details

- {{< fa calendar >}} &nbsp; {{< var course.days >}}
- {{< fa calendar-alt >}} &nbsp; {{< var course.dates >}}
- {{< fa clock >}} &nbsp; {{< var course.time1 >}}
- &nbsp; {{< var course.time2 >}}
- {{< fa location-dot >}} &nbsp; {{< var course.location >}}
:::

::: {.g-col-6}
### Instructor

- {{< fa user >}} &nbsp; [{{< var instructor.name >}}]({{< var instructor.url >}})
- {{< fa university >}} &nbsp; {{< var instructor.office >}}
- {{< fa envelope >}} &nbsp; [{{< var instructor.email >}}](mailto:{{< var instructor.email >}})
- {{< fa university >}} &nbsp; Office hours: [M 9-11 am and Th 3:30-5:30 pm]({{< var instructor.appointment_url >}}) and by appointment
:::

:::



# Syllabus

You can find a PDF version of the syllabus with the weekly schedule on the sidebar of this page.

:::


## Course Overview

This course discusses the ways in which religion and politics intersect in the United States. The topics that will be covered include the relationship between church and state; the connection between an individual's religion (i.e., their beliefs and behavior) and their political attitudes and behaviors; how political candidates make use of religion; churches as political communities; the rise of the "Religious Right" and Christian nationalism; and the intersection of religion and politics for racial, ethnic, and religious minorities. The course will highlight important concepts and theories as well as the measures and data that scholars use to study the influence of religion on American politics.


## Learning Outcomes

- Develop an understanding of the role that America's religious traditions have played in contemporary politics.
- Think critically about the ways in which religion influences politics and politics influences religion.
- Examine the intersection of religion and politics through a social scientific perspective, generating questions and ways to answer them.
- Apply the theories and concepts from religion and politics research to current events.


## Work and Assessment


### Required Text:

- Putnam, Robert and David Campbell. 2010. *American Grace.* New York: Simon & Schuster. {{< fa book >}}

    All other required readings are linked below or are stored in a Dropbox folder ({{< fa brands dropbox title="Available on Dropbox folder" >}}). I will provide the link to the folder via Canvas.



### Course Expectations

Students are expected to attend class and participate during class discussions. This course has both a lecture and a discussion component. The lectures will incorporate the readings assigned for that day/week. I expected you will have completed all readings prior to class and will come prepared with comments and questions. Readings can be accessed through the library resources or, when not available through the library, via Canvas.

- <ins>*A note of caution:*</ins> many, if not most, of the readings come from academic journals and books, mostly from political science but also from sociology. You will find that many of the articles are quite technical and they can be hard to follow even when one is familiar with the methodology. Do not despair. Make sure to understand, and wrestle with, the main ideas presented in a piece. Pay attention as well to how the authors support their arguments with data without getting bogged down by technical details.

### Course Workload

As a course with a course count of one, you are expected to spend 8-10 hours of work each week (not including our two 75-minute weekly course meetings).


### Assignments and grades

Final course grades will be assigned as follows:

```{r grades, include=FALSE}
grading <- tribble(
  ~Grade, ~Range, ~Grade1, ~Range1,
  "A",  "94–100%", "C",  "74–76%",
  "A−", "90–93%", "C−", "70–73%",
  "B+", "87–89%", "D+", "67–69%",
  "B",  "84–86%", "D",  "64–66%",
  "B−", "80–83%", "D−", "60–63%",
  "C+", "77–79%", "F",  "< 60%"
)
```

<div class="centered-table">

```{r show-grades-table, echo=FALSE, results="asis"}
grading %>%
  pandoc.table(justify = "lclc",
               col.names = c("Grade", "Range", "Grade", "Range"))
```

<div>


Final grades will be calculated based on the following components:

**Participation (10%)**: All students are expected to be active participants in class discussions. Our learning experience will be enhanced if you come prepared and ready to share your comments, insights, and/or questions from the material covered in the lectures and readings with the rest of the class.

  - <ins>*Attendance policy:*</ins> Attendance is part of your participation grade, and while there is no separate grade for attendance it is a prerequisite for in-class participation. That said, to accommodate unforeseeable circumstances, you have two *unexcused* absences that will not affect your participation grade. Any additional absences will hurt your participation grade. If you accumulate four *unexcused* absences in the semester, you will forfeit your entire participation grade. Note that students who are absent, whether excused or not, are still expected to submit the weekly summary corresponding to that class period. (Absences due to extenuating circumstances will be considered separately. Please refer to the [College's Class Attendance Policy](https://catalog.holycross.edu/requirements-policies/academic-policies/#coursepoliciestext-otp2) for the college-wide policy on absences and be in communication if a situation that will keep you from attending class for an extended period of time arises.)





**Weekly summaries (15%)**: Each week for 10 weeks, you will summarize one article from the week's readings (the reading assgined for a summary has a {{< fa pen >}} after it in the course schedule). There are 13 weeks in the semester where you can write a summary, so you can skip 3 weeks. It is up to you to decide which weeks to skip. The summary must be **no longer** than three-quarters of a page (single-spaced, 12-point font). These summaries must be submitted on Canvas before class time *on the day the reading was assigned*. Please see the [guide](resource/summaries.qmd) on the course website for the format of these summaries.


**Congregational visit paper (30% total)**: You will write an empirical research paper (4-6 pages, double-spaced) similar to the ones political scientists publish in academic journals. For this assignment, you will attend at least one religious service of a religious tradition or denomination other than your own and will write about how politics is (or is not) addressed, directly or indirectly, in that particular congregation. The paper will be broken down into the following parts that you will submit throughout the semester:

  1. <ins>*Front end of paper (2-3 pages, 7.5% of final grade):*</ins> This portion of the paper includes the motivation, research questions, literature review, theory, expectations, and methodology. This will be due early in the semester for you to do your best and get feedback.
  
  2. <ins>*Full draft for peer review (4-6 pages, 7.5% of final grade):*</ins> After collecting data for the paper, you will submit a full draft of the paper for peer review. Another student will provide feedback on your work and you will do the same for them.
  
  3. <ins>*Final draft (4-6 pages, 15% of final grade):*</ins> After the peer review, you will submit the final draft of the paper. I expect you will have included the feedback you received on the first portion of the paper as well as the feedback from your classmate.

**Op-ed (15%)**: You will write an op-ed (600-1,200 words) on a course-related topic of your choice. Students must make a clear argument and use class material to help them organize their essays or provide context to the situation.

- Papers must be submitted via Canvas. Late submissions will be penalized: half a point (out of 10) if turned in late but within 24 hours and one point for each 24-hour period after the due date.


**Midterm and Final Exam (15% each)**: There will be a midterm and final exam where you will answer three medium-length questions. Both the midterm and final exams will be in-person. These exams will assess your familiarity with the terms and theories in the religion and politics literature that we will have covered in class.



## Course Conduct

A common saying advises never to discuss politics or religion in polite company. In this class, we will discuss politics *and* religion and we will do so in a respectful environment. As is bound to be the case when we talk about religion, we will discuss many topics that might be controversial or of a sensitive nature for some or all in our group. Our goal in this class is to have constructive discussions about how religion and politics intersect. Derogatory comments toward other students or to members of a particular religious group will not be tolerated.


## Communication Policy

If you have questions about class in general or about a specific assignment, you can email me or come by my office. You can expect a response within 24 hours (unless you email me on Friday). Students are also more than welcome to attend office hours, as this is time set aside specifically for you (really, they should be called "student hours"). In addition, if my office door is open, feel free to stop by. Please note that I will not be "pre-grading" assignments, but I am happy to look over your work together during office hours.


## Regrading Policy

If a student wishes to appeal a grade, they must do so in writing within one week of receiving their graded assignment back. The student must indicate clearly what part of the assignment's grading they disagree with, making sure to respond to the instructor's comments in that section and making their case for why they should not have been deducted points. These requests should be done via email and the subject name should read "POLS 299-F06 Regrading request - [Name of assignment]." 


## Academic Integrity and Generative AI Policy

All education is a cooperative enterprise between faculty and students. This cooperation requires trust and mutual respect, which are only possible in an environment governed by the principles of academic integrity. As an institution devoted to teaching, learning, and intellectual inquiry, Holy Cross expects all members of the College community to abide by the highest standards of academic integrity. Any violation of academic integrity undermines the student-faculty relationship, thereby wounding the whole community. Students in this class are required to read the full text of the [College’s Academic Integrity Policy](https://catalog.holycross.edu/requirements-policies/academic-policies/#academicintegritytext) and to abide by its standards.


Unless otherwise noted, you are expected to complete all coursework independently without the use of generative AI tools (e.g., ChatGPT, Claude, Gemini). Because this course emphasizes using writing as a tool for learning and a vehicle for developing a unique writerly voice, using AI tools runs counter to our goals for writing assignments. Because the assumption in this course is that all writing students submit is original to them, cases where students turn in AI-generated content will be considered cheating.




## Course Accessibility and Academic Accommodations


The College of the Holy Cross is committed ["to providing students with disabilities equal access to the educational opportunities and programs available at Holy Cross in accordance with the Americans with Disabilities Act of 1990 as amended, Section 504 of the Rehabilitation Act of 1973, and Massachusetts laws."](https://catalog.holycross.edu/requirements-policies/academic-policies/#coursepoliciestext)

Any student who feels the need for accommodation based on the impact of a disability should contact the [Office of Accessibility Services](https://www.holycross.edu/community/health-wellness/accessibility-accommodations) to discuss support services available. Once the office receives documentation supporting the request for accommodation, the student would meet privately with Accessibility Services to discuss reasonable and appropriate accommodations.

The office can be reached by calling 508-793-3693. If you are already registered with Accessibility Services, please be sure to get your accommodation letters and deliver them to your instructors in a timely fashion. Instructors need at least 4-5 days advance notice to be able to facilitate the process of receiving testing accommodations.



## Wellbeing and Mental Health

The wellbeing and mental health of students is important. If you are having trouble managing your coursework and are experiencing unusual amounts of stress, sadness, or anxiety, the College has resources you may find useful. Please reach out to [Student Wellness Education](https://www.holycross.edu/community/health-wellness/student-wellness-education) or the office of [Counseling and Psychological Services](https://www.holycross.edu/community/health-wellness/counseling-psychological-services).


## Title IX

If you have experienced any form of sexual harassment or violence, dating or domestic violence, or stalking, please talk to any faculty or staff member with whom you feel comfortable. Faculty and staff members want to support you and have been trained to help. They will also inform the Title IX office so that you learn about options available to you. If you do not want college administrators to know what you have experienced, you may speak to a chaplain or counseling services with complete confidentiality. They will not tell anyone what you share with them unless you give your express permission. You may contact the Title IX Office directly at <titleix@holycross.edu>.


## Chosen Names and Personal Pronouns
 
Everyone has the right to be addressed by the name and pronouns that correspond to their gender identity. Rosters do not list gender or pronouns, so you may indicate the pronouns you use so that I don't make assumptions based on your name and/or appearance/self-presentation. If you use a chosen name, please let me know. Chosen names and pronouns are to be respected at all times in the classroom. Mistakes in addressing one another may happen, so I encourage an environment of openness to correction and learning. I will not however, tolerate repeated comments which disrespect or antagonize students who have indicated pronouns or a chosen name.


## Center for Writing and Writer's Workshop

The Center for Writing offers one-on-one appointments with trained peer consultants trough their Writer's Workshop who can help you improve your communication skills. You are encouraged to use this resource when working on your class assignments to make your arguments more compelling, your sentences clearer, and persistent grammatical errors nonexistent. You can schedule an appointment at <https://holycross.mywconline.com/>.



## Recording of Classes

Please note that, consistent with applicable federal and state law, this course may be video/audio recorded as an accommodation *only* with permission from the Office of Accessibility Services. Students are not permitted to record the contents of this class under any other circumstances.



## Syllabus Change Policy

This syllabus is only a guide for the course and is subject to change with advanced notice.


::: {.content-visible when-format="pdf"}
  
## Schedule

{{< include content/01-week1.qmd >}}

{{< include content/02-week2.qmd >}}

{{< include content/03-week3.qmd >}}

{{< include content/04-week4.qmd >}}

{{< include content/05-week5.qmd >}}

{{< include content/06-week6.qmd >}}

{{< include content/07-week7.qmd >}}

{{< include content/08-week8.qmd >}}

{{< include content/09-week9.qmd >}}

{{< include content/10-week10.qmd >}}

{{< include content/11-week11.qmd >}}

{{< include content/12-week12.qmd >}}

{{< include content/13-week13.qmd >}}

{{< include content/14-week14.qmd >}}

{{< include content/15-week15.qmd >}}

{{< include content/16-week16.qmd >}}

:::

{{< include index_title.qmd >}}
